Estimating Information Transfer in Hard Parameter Sharing via Precise Bias-variance Asymptotics

### 2021-07-17 revision_journal/multitask.tex
wip
finish sec 5
finish sec 4 figures
finish sec 4

\paragraph{Data model.} Suppose we have $t$ datasets whose sample sizes are all equal to $n$ and whose feature covariates are all equal to $X \in \real^{n \times p}$. The label vector of the $i$-th task follows a linear model
\begin{align}\label{eq_mtl_data}
    Y^{(i)} = X \beta^{(i)} + \varepsilon^{(i)}, \text{ for } i=1, 2,\cdots, t.
\end{align}
Similar to Section \ref{sec_main}, we use the first $(t-1)$ datasets as sources to predict the $t$-th task.
However, there is a \text{model shift} between the data sources and the task we would like to predict.
We make several standard assumptions on $X$ and each of $\varepsilon^{(i)}$.
First, $X = Z\Sigma^{1/2}$ is a random matrix satisfying Assumption \ref{assm_big1} (same as $X^{(2)}$).
In particular, the sample size $n$ is greater than the dimension $p$.
Second, every $\varepsilon^{(i)} \in \real^{n}$ is a random vector with i.i.d entries of mean zero, variance $\sigma^2$, and bounded moments up to any order (cf. equation \eqref{eq_highmoments}).
Finally, $\beta^{(i)} \in \real^{p}$ is a fixed vector independent from any other $\beta^{(j)}$ for $j \neq i$, the matrix $X$, and $\varepsilon^{(j)}$ for any $j$.

@@

Let $(\hat A, \hat B)$ denote a global minimizer of $f(A, B)$.
We define the HPS estimator for task $t$ as $\hat \beta_t^{\MTL} := \hat B \hat A_i$, where $\hat A_i$ denotes the $i$-th column of $\hat A$.
We evaluate the performance of $\hat{\beta}_t^{\MTL}$ according to its excess risk:
\be\label{ith_loss}
    L_t(\hat{\beta}_t^{\MTL}) = \bignorm{\Sigma^{1/2} \left(\hat{\beta}^{\MTL}_t - \beta^{(t)}\right)}^2.
\ee

\paragraph{Result.} We show that in the multi-task setting, hard parameter sharing finds the ``best'' rank-$r$ approximation to all tasks.
To formally describe our result, we introduce several notations.
Let $B^\star \define [{\beta}^{(1)},{\beta}^{(2)},\dots,{\beta}^{(t)}] \in \real^{p\times t}$ be the concatenated model vectors of all tasks.
Let $A^{\star} {A^{\star}}^{\top}$ be the best rank-$r$ approximation of ${B^{\star}}^\top\Sigma B^{\star}$ in the following sense:
\begin{align}\label{eq_A_star}
	A^{\star} \define \argmin_{U\in\real^{t\times r} : U^{\top} U = \id_{r\times r}} \inner{U U^{\top}} {{B^{\star}}^{\top} \Sigma B^{\star}},
\end{align}
where $\langle \cdot ,\cdot \rangle $ denotes the Frobenius inner product between two matrices.
Let $a_i^{\star}\in\real^r$ be the $i$-th column vector of $A^{\star}{A^{\star}}^{\top}$.
We characterize the excess risk $L_t(\hat{\beta}_t^{\MTL})$ precisely in the following result.

\begin{theorem}[Excess risk of HPS for multiple tasks under model shift]\label{thm_many_tasks}
Suppose the multi-task setting according to equation \eqref{eq_mtl_data} holds.
Let $r$ be any positive integer less than $t$.
Suppose the $r$-th largest eigenvalue of ${B^\star}^\top \Sigma B^\star$ is strictly larger than its $(r+1)$-th largest eigenvalue. 
Let $c>0$ be an arbitrary (small) constant.
The following estimates for $L_t(\hat{\beta}_t^{\MTL})$ holds with high probability:
\begin{align}
	\bigabs{L_t(\hat{\beta}_t^{\MTL}) - L_t(B^{\star}a_t^{\star}) -\frac{p\sigma^2}{n-p}  \norm{a_t^{\star}}^2  } 
	\le  \sqrt{\bignorm{{B^\star}^\top\Sigma B^\star}_{\op} \cdot n^{-\frac 1 2+ \frac 2{\varphi} + c}  + \sigma^2 n^{-\frac 1 2 + c}} \cdot \frac{\bignorm{{B^\star}^\top\Sigma B^\star}_{\op}+  \sigma^2}{\lambda_r - \lambda_{r+1}}. \label{Li_multi1}
\end{align}
	%where $\lambda_r({B^\star}^\top\Sigma B^\star)$ and $\lambda_{r+1}({B^\star}^\top\Sigma B^\star)$ are respectively be the $r$-th and $(r+1)$-th largest eigenvalue of ${B^\star}^\top\Sigma B^\star$.
%	where $C_1: = \frac{\bignormFro{\Sigma^{1/2} B^{\star}}^2 + \sigma^2 t}{\lambda_r ({B^\star}^\top \Sigma B^\star)- \lambda_{r+1}({B^\star}^\top \Sigma B^\star)}$. %and $C_2 :=  C_1\cdot \norm{\Sigma^{1/2} B^{\star}}$.
%Finally, we have a better bound for the averaged prediction loss:  with high probability,
%\begin{align}
%&\left|\frac{1}{t}\sum_{i=1}^t L_i(\hat{\beta}_i^{\MTL}) - \frac1t\bignorm{\Sigma^{1/2} B^{\star} (A^\star {A^\star}^{\top} - \id_{t\times t})}_F^2 - \frac{p \sigma^2}{n-p}\cdot \frac{r}{t}  \right| \nonumber \\
% &\le n^{-1/2+2/\varphi+c}  \norm{{B^\star}^\top\Sigma B^\star}+ n^{-1/2+c}   \sigma^2 .\label{Li_multi2}
%\end{align}
\end{theorem}

Recall that $\varphi > 4$ according to Assumption \ref{assm_big1}, thus, $\frac {2}{\varphi} \le \frac 1 2$ and $n^{-\frac 1 2 + \frac 2 {\varphi} + c}$ vanishes to zero for a small enough constant $c$.
Therefore, equation \eqref{Li_multi1} precisely characterizes the excess risk of the HPS estimator.
As a remark, the same result applies to any other tasks although we have focused on predicting task $t$.
Theorem \ref{thm_many_tasks} significantly extends the two-task model shift setting of Theorem \ref{cor_MTL_loss}.
The term $L_t(B^{\star} a_t^{\star})$ captures the bias of the HPS estimator and $\frac{p \sigma^2}{n - p} \bignorm{a_t^{\star}}^2$ captures the estimator's variance.
%The bound \eqref{minimizer_beta} verifies our intuition that hard parameter sharing approximates the matrix ${B^{\star}}^\top\Sigma B^{\star}$ through a best rank-$r$ subspace. The estimate \eqref{Li_multi1}
% gives the exact asymptotic limit of $L_i(\hat{\beta}_i^{\MTL}(\hat a))$, which shows that the prediction loss of $\hat{\beta}_i^{\MTL}$ decomposes into a bias term $L_i(B^{\star} a_i^{\star})$ that measures the prediction loss of $B^{\star} a_i^{\star}$, plus a variance term that scales with $\norm{a_i^{\star}}^2$. Since $\norm{a_i^{\star}}^2\le 1$, compared with the single-task predication loss \eqref{L_STL_simple01}, the variance term always decreases for the multi-task HPS estimator. On the other hand, the bias term always increases (because the bias in single-task linear regression is zero). Hence as in the transfer learning setting, whether the HPS estimator is better than the OLS estimator depends on an intricate  {bias-variance tradeoff}.
%{\cor One direct implication of our result is that compared to STL, the variance always decreases, since STL's variance is equal to $\sigma^2 \tr[\Sigma (X^{\top} X)^{-1}]$.
%On the other hand, the bias always increases.}
%We can observe a similar {bias-variance tradeoff} for the averaged predication loss in \eqref{Li_multi2} using the fact that $r<t$. Note that the estimate \eqref{Li_multi2} can be applied even when the best rank-$r$ subspace approximation of ${B^{\star}}^\top\Sigma B^{\star}$ is not unique. For all the estimates in Theorem \ref{thm_many_tasks}, we believe that their convergence rates are asymptotically tight when $n$ goes to infinity.
%	For the optimization objective  in \eqref{eq_mtl_same_cov}, using the local optimality condition over $B$, that is, $\frac{\partial f}{\partial B} = 0$, we obtain $\hat{B}$ as a function of $A$:
%	\begin{align}
%		\hat{B}(A) %&\define (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{j=1}^t Y^{(j)} A_j^{\top}} (A  A^{\top})^{+} \nonumber \\
%		&= (X^{\top} X)^{-1} X^{\top} Y A^{\top} (AA^{\top})^{+}, \label{eq_Bhat}
%	\end{align}
%	where $Y := [Y^{(1)}, Y^{(2)}, \dots, Y^{(t)}]$ and $(AA^{\top})^{+}$ denotes the pseudoinverse of $AA^{\top}$.
	%Here we have used that $X^{\top}X$ is invertible since $n > \rho \cdot p$ and $\rho > 1$ (cf. Fact \ref{fact_minv}).
	%\FY{Is $\dag$ a standard notation? It is a bad notation at least for me because $\dag$ is more often used as Hermitian conjugate. Wiki page uses $(AA^{\top})^{+}$ for pseudo-inverse.}
%	Plugging $\hat{B}(A)$ into equation \eqref{eq_mtl_same_cov}, we obtain the following objective that only depends on $A$ (in matrix notation):
%	\begin{align}\label{eq_mtl_output_layer}
%		g(A) = \bignormFro{X (X^{\top}X)^{-1}X^{\top} Y A^{\top} (AA^{\top})^{+} A - Y}^2.
%	\end{align}
The proof of Theorem \ref{thm_many_tasks}, which is based on a characterization of the global minimum of problem \eqref{eq_mtl_same_cov}, can be found in Section  \ref{app_proof_error_same_cov}.

\begin{figure*}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/same_covariates.eps}
		\caption{Example \ref{ex_same_cov}}
		\label{fig_same_cov}
	\end{subfigure}\hfill
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.9\textwidth]{figures/same_covariates.eps}
		\caption{Example \ref{ex_sample_ratio}}
		\label{fig_width}
	\end{subfigure}
%	\caption{%Three takeaways of our theory in Section \ref{sec_insight}.
%	Our estimated losses (solid line) match the empirical losses (dotted line) accurately under various settings in dimension $p = 200$.
%	\textbf{Left.} Validating Example \ref{ex_same_cov} for ten tasks: the noise variance $\sigma^2$ is $1/4$.
%	\textbf{Middle.} Validating Example \ref{ex_sample_ratio} for two tasks: we discover an interesting phenomena by fixing task two's sample size and increasing task one's sample size.
%	Moreover, our result accurately predicts the critical point (marked in circle) of the loss curve.
%	Depending on how large the distance $d^2$ is, task two's prediction loss decreases initially before increasing again, or decreases monotonically.
%	\textbf{Right.} We show how different levels of covariate shift affect hard parameter sharing when there is no bias.
%	Having covariate shift increases task two's prediction loss when task two's sample size is smaller than task one. Otherwise, having covariate shift (surprisingly) decreases task two's prediction loss.}
%	\label{fig_model_shift_phasetrans}
\end{figure*}

@@

\begin{example}[How to set the width $r$ in HPS?]
\label{ex_same_cov}
%Suppose every $\beta^{(i)}$ consists of two random components, one that is shared among all tasks and one that is task-specific.
%Thus, each task contributes a certain amount to the shared component and injects a task-specific bias.
%More precisely, %we have
An important question in applications of HPS is how to set the width $r$ of the feature layer.
We use our estimates to show that the optimal width depends on the tradeoff between the estimator's bias and variance.
Concretely, consider a random-effect model where all tasks share a model vector $\beta_0$ plus perturbations.
For every $i=1,2,\cdots, t,$, $\beta^{(i)} = \beta_0 + \wt \beta^{(i)}$, where $\wt \beta^{(i)}$ denotes the $i$-th task-specific component whose entries are i.i.d. Gaussian random variables of mean zero and variance $\frac {\kappa^2} p$.

\begin{claim}
\begin{enumerate}
	\item {\bf Positive vs. negative transfer.} The averaged HPS prediction loss is smaller than the single-task OLS prediction loss if and only if $\frac{d^2}{p}\tr \Sigma  < \frac{p\sigma^2 }{n - p}$, that is, the ``task-specific variance'' is smaller than the ``noise variance'' up to some constant factor.

	\item {\bf The optimal rank $r$.} If $\frac{d^2}{p}\tr \Sigma  < \frac{\sigma^2 p}{n - p}$, then the smallest averaged HPS  prediction loss is achieved when $r=1$. Hence increasing the width $r$ of the shared feature representation layer does not help.

@@

    In this random-effect model, using the concentration of Gaussian random vectors (e.g. Lemma \ref{largedeviation} in the supplement  \cite{MTL_suppl}),
    %Based on the definition of the random-effect model,
    the $(i, j)$-th entry of ${B^{\star}}^{\top} \Sigma B^{\star}$ is equal to %(ignoring lower order terms)
    %\begin{align*}
    %	\beta_i^{\top} \beta_j \approx \norm{\beta_0}^2 + \frac{d^2}{2}\delta_{ij},\quad 1\le i,j \le t,
    %\end{align*}

@@

    	\beta_i^{\top}\Sigma  \beta_j =\beta_0^\top \Sigma \beta_0 + \delta_{ij} \frac{d^2 }{p}\tr \Sigma + \OO\left(p^{-1/2+c}\|\beta_0\|^2+ p^{-1/2+c} d^2\right),

@@

     with high probability for any constant $c>0$. We omit the details to show the error bound using Lemma \ref{largedeviation}.
    %Note that $\norm{\beta_0}^2$ is approximately $\kappa^2$.
    With \eqref{betaSbeta}, it is easy to calculate that with high probability, the eigenvalues of ${B^{\star}}^{\top} \Sigma B^{\star}$ are given by
    $$\lambda_1=\left[1+\OO(p^{-1/2+c})\right]\cdot\left(t \beta_0^\top \Sigma \beta_0  + \frac{d^2}{ p}\tr \Sigma\right) ,$$
    and
    $$ \lambda_i=\left[1+\OO(p^{-1/2+c})\right]\cdot \frac{d^2}{ p}\tr \Sigma , \ \ i=2,\cdots, t.$$ %Therefore, by taking a rank-$1$ approximation of ${B^{\star}}^{\top} B^{\star}$, we get the average prediction loss of $B^{\star} a_i^{\star}$.
    Thus for the best rank-$r$ approximation $A^\star {A^\star}^\top$  of ${B^{\star}}^{\top}\Sigma  B^{\star}$, we have
    $$\bignorm{\Sigma^{1/2} B^{\star} (A^\star {A^\star}^{\top} - \id_{t\times t})}_F^2= [1+\OO(p^{-1/2+c})]\cdot(t-r)\frac{d^2}{ p}\tr \Sigma $$
    with high probability.

@@

    Then using \eqref{Li_multi2}, we obtain that
    \[ \frac{1}{t}\sum_{i=1}^t L_i(\hat{\beta}_i^{\MTL}) = \bigbrace{1 - \frac{r}{t}} \frac{d^2}{p}\tr \Sigma + \frac{r}{t} \cdot \frac{p\sigma^2 }{n - p} +\oo(\|\beta_0\|^2+d^2 + \sigma^2) ,\quad \text{w.h.p. }\]
    %We describe a proof sketch.
    %First, we show that the bias equation $L(B^{\star} a_i^{\star})$ simplifies to the following
    %\[ \frac{1}{t} \sum_{i=1}^t L(B^{\star} a_i^{\star}) = \frac{1}{t}\normFro{B^{\star} A^{\star} {A^{\star}}^{\top} - B^{\star}}^2 \approx \left(1 - \frac{1}{t}\right) {d^2}{}  . \]
    %Second, using Fact \ref{fact_tr}, one can see that the average variance is
    %\begin{align*}
    %	\frac{1}{t} \sum_{i=1}^t \sigma^2\norm{a_i^{\star}}^2 \bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{\sigma^2}{t} \sum_{i=1}^t \norm{a_i^{\star}}^2 \frac{p}{n - p}
    %	= \frac{1}{t}\frac{\sigma^2 p}{n - p},
    %\end{align*}
    %because $A^{\star}$ has rank-$1$ and $\sum_{i=1}^t \norm{a_i^{\star}}^2 = 1$.
    %Combined together, we have derived the average prediction loss in the random-effect model.
    %If the error is sufficiently small, then
    %Recall that the average prediction loss of STL scales as $\sigma^2\cdot \bigtr{\Sigma (X^{\top} X)^{-1}} = \frac{\sigma^2 p}{n - p}$ by Fact \ref{fact_tr}.
    %Comparing HPS to STL, we have the following qualitative properties.
    %	Suppose $n$ is sufficiently large so that the error is negligible.

@@

\end{example}

%We demonstrate the accuracy of our results in simulations.
%While our theory is asymptotic (with error terms that are negligible when $p$ is sufficiently large), we observe that they are incredibly accurate in a moderate dimension of $p = 200$.

>>>>

\paragraph{Data model.} Suppose we have $t$ datasets whose sample sizes are all equal to $n$ and whose feature covariates are all equal to $X \in \real^{n \times p}$. The label vector of the $i$-th task follows a linear model
\begin{align}\label{eq_mtl_data}
    Y^{(i)} = X \beta^{(i)} + \varepsilon^{(i)}, \text{ for } i=1, 2,\cdots, t.
\end{align}
Similar to Section \ref{sec_main}, we use the first $(t-1)$ datasets as sources to predict the $t$-th task.
However, there is a \text{model shift} between the data sources and the task we would like to predict.
We make several standard assumptions on $X$ and each of $\varepsilon^{(i)}$.
First, $X = Z\Sigma^{1/2}$ is a random matrix satisfying Assumption \ref{assm_big1} (same as $X^{(2)}$).
In particular, the sample size $n$ is greater than the dimension $p$.
Second, every $\varepsilon^{(i)} \in \real^{n}$ is a random vector with i.i.d entries of mean zero, variance $\sigma^2$, and bounded moments up to any order (cf. equation \eqref{eq_highmoments}).
Finally, $\beta^{(i)} \in \real^{p}$ is a fixed vector independent from any other $\beta^{(j)}$ for $j \neq i$, the matrix $X$, and $\varepsilon^{(j)}$ for any $j=1,\dots,t$.

@@

Let $(\hat A, \hat B)$ denote a global minimizer of $f(A, B)$.
We define the HPS estimator for task $i$ as $\hat \beta_i^{\MTL} := \hat B \hat A_i$, where $\hat A_i$ denotes the $i$-th column of $\hat A$.
We evaluate the performance of $\hat{\beta}_i^{\MTL}$ according to its excess risk:
\be\label{ith_loss}
    L_i(\hat{\beta}_i^{\MTL}) = \bignorm{\Sigma^{1/2} \left(\hat{\beta}^{\MTL}_i - \beta^{(i)}\right)}^2.
\ee

\paragraph{Result.} We show that in the multi-task setting, hard parameter sharing finds the ``best'' rank-$r$ approximation to all tasks.
To formally describe our result, we introduce several notations.
Let $B^\star \define [{\beta}^{(1)},{\beta}^{(2)},\dots,{\beta}^{(t)}] \in \real^{p\times t}$ be the concatenated model vectors of all tasks.
Let $A^{\star} {A^{\star}}^{\top}$ be the best approximation of ${B^{\star}}^\top\Sigma B^{\star}$ in the set of rank-$r$ subspaces:
\begin{align}\label{eq_A_star}
	A^{\star} \define \argmax{U\in\real^{t\times r} : U^{\top} U = \id_{r\times r}} \inner{U U^{\top}} {{B^{\star}}^{\top} \Sigma B^{\star}},
\end{align}
where $\langle \cdot ,\cdot \rangle $ denotes the Frobenius inner product between two matrices.
Let $a_i^{\star}\in\real^r$ be the $i$-th column vector of $A^{\star}{A^{\star}}^{\top}$.
We show a precise estimate of $L_i(\hat{\beta}_i^{\MTL})$ based on $a_i^{\star}$ as follows.

\begin{theorem}[Excess risk of HPS for multiple tasks under model shift]\label{thm_many_tasks}
Suppose the multi-task setting according to equation \eqref{eq_mtl_data} holds.
Let $r < t$ be a positive integer.
Suppose the $r$-th largest eigenvalue of ${B^\star}^\top \Sigma B^\star$ is strictly larger than its $(r+1)$-th largest eigenvalue. 
Let $c>0$ be an arbitrarily (small) constant.
The following estimate of $L_i(\hat{\beta}_i^{\MTL})$ holds w.h.p., for any $i = 1,\dots,t$:
\begin{align}
	\bigabs{L_i(\hat{\beta}_i^{\MTL}) - L_i(B^{\star}a_i^{\star}) -\frac{p\sigma^2}{n-p} \cdot \norm{a_i^{\star}}^2} 
	\le  \sqrt{\norm{{B^\star}^\top\Sigma B^\star}_{2} \cdot n^{-\frac 1 2+ \frac 2{\varphi} + c}  + \sigma^2 n^{-\frac 1 2 + c}} \cdot \frac{\norm{{B^\star}^\top\Sigma B^\star}_{2}+  \sigma^2}{\lambda_r - \lambda_{r+1}}. \label{Li_multi1}
\end{align}
	%where $\lambda_r({B^\star}^\top\Sigma B^\star)$ and $\lambda_{r+1}({B^\star}^\top\Sigma B^\star)$ are respectively be the $r$-th and $(r+1)$-th largest eigenvalue of ${B^\star}^\top\Sigma B^\star$.
%	where $C_1: = \frac{\bignormFro{\Sigma^{1/2} B^{\star}}^2 + \sigma^2 t}{\lambda_r ({B^\star}^\top \Sigma B^\star)- \lambda_{r+1}({B^\star}^\top \Sigma B^\star)}$. %and $C_2 :=  C_1\cdot \norm{\Sigma^{1/2} B^{\star}}$.
%Finally, we have a better bound for the averaged prediction loss:  with high probability,
%\begin{align}
%&\left|\frac{1}{t}\sum_{i=1}^t L_i(\hat{\beta}_i^{\MTL}) - \frac1t\bignorm{\Sigma^{1/2} B^{\star} (A^\star {A^\star}^{\top} - \id_{t\times t})}_F^2 - \frac{p \sigma^2}{n-p}\cdot \frac{r}{t}  \right| \nonumber \\
% &\le n^{-1/2+2/\varphi+c}  \norm{{B^\star}^\top\Sigma B^\star}+ n^{-1/2+c}   \sigma^2 .\label{Li_multi2}
%\end{align}
\end{theorem}

Recall that $\varphi > 4$ according to Assumption \ref{assm_big1}, thus, $\frac {2}{\varphi} \le \frac 1 2$ and $n^{-\frac 1 2 + \frac 2 {\varphi} + c}$ vanishes to zero for a small enough constant $c$.
Thus, $L_i(B^{\star} a_i^{\star}) + \frac{p\sigma^2}{n - p} \norm{a_i^{\star}}^2$ is the limit of $L_i(\hat{\beta}_i^{\MTL})$ as $p$ approaches infinity---$L_i(B^{\star} a_i^{\star})$ is the limiting bias of $\hat{\beta}_i^{\MTL}$ and $\frac{p \sigma^2}{n - p} \bignorm{a_i^{\star}}^2$ is its limiting variance.
The proof of Theorem \ref{thm_many_tasks}, which characterizes the global minimum of problem \eqref{eq_mtl_same_cov} through a connection to PCA, can be found in Section \ref{app_proof_error_same_cov}.
%We also point out that equation \eqref{Li_multi1} similarly applies to any other task $t' \neq t$.
%The bound \eqref{minimizer_beta} verifies our intuition that hard parameter sharing approximates the matrix ${B^{\star}}^\top\Sigma B^{\star}$ through a best rank-$r$ subspace. The estimate \eqref{Li_multi1}
% gives the exact asymptotic limit of $L_i(\hat{\beta}_i^{\MTL}(\hat a))$, which shows that the prediction loss of $\hat{\beta}_i^{\MTL}$ decomposes into a bias term $L_i(B^{\star} a_i^{\star})$ that measures the prediction loss of $B^{\star} a_i^{\star}$, plus a variance term that scales with $\norm{a_i^{\star}}^2$. Since $\norm{a_i^{\star}}^2\le 1$, compared with the single-task predication loss \eqref{L_STL_simple01}, the variance term always decreases for the multi-task HPS estimator. On the other hand, the bias term always increases (because the bias in single-task linear regression is zero). Hence as in the transfer learning setting, whether the HPS estimator is better than the OLS estimator depends on an intricate  {bias-variance tradeoff}.
%{\cor One direct implication of our result is that compared to STL, the variance always decreases, since STL's variance is equal to $\sigma^2 \tr[\Sigma (X^{\top} X)^{-1}]$.
%On the other hand, the bias always increases.}
%We can observe a similar {bias-variance tradeoff} for the averaged predication loss in \eqref{Li_multi2} using the fact that $r<t$. Note that the estimate \eqref{Li_multi2} can be applied even when the best rank-$r$ subspace approximation of ${B^{\star}}^\top\Sigma B^{\star}$ is not unique. For all the estimates in Theorem \ref{thm_many_tasks}, we believe that their convergence rates are asymptotically tight when $n$ goes to infinity.
%	For the optimization objective  in \eqref{eq_mtl_same_cov}, using the local optimality condition over $B$, that is, $\frac{\partial f}{\partial B} = 0$, we obtain $\hat{B}$ as a function of $A$:
%	\begin{align}
%		\hat{B}(A) %&\define (X^{\top}X)^{-1} X^{\top} \bigbrace{\sum_{j=1}^t Y^{(j)} A_j^{\top}} (A  A^{\top})^{+} \nonumber \\
%		&= (X^{\top} X)^{-1} X^{\top} Y A^{\top} (AA^{\top})^{+}, \label{eq_Bhat}
%	\end{align}
%	where $Y := [Y^{(1)}, Y^{(2)}, \dots, Y^{(t)}]$ and $(AA^{\top})^{+}$ denotes the pseudoinverse of $AA^{\top}$.
	%Here we have used that $X^{\top}X$ is invertible since $n > \rho \cdot p$ and $\rho > 1$ (cf. Fact \ref{fact_minv}).
	%\FY{Is $\dag$ a standard notation? It is a bad notation at least for me because $\dag$ is more often used as Hermitian conjugate. Wiki page uses $(AA^{\top})^{+}$ for pseudo-inverse.}
%	Plugging $\hat{B}(A)$ into equation \eqref{eq_mtl_same_cov}, we obtain the following objective that only depends on $A$ (in matrix notation):
%	\begin{align}\label{eq_mtl_output_layer}
%		g(A) = \bignormFro{X (X^{\top}X)^{-1}X^{\top} Y A^{\top} (AA^{\top})^{+} A - Y}^2.
%	\end{align}

@@

\paragraph{Illustrative examples.}
We give an example of our estimate in the random-effect model.
In this setting, the model vector of every task is the sum of a shared vector $\beta_0$ plus a task-specific component $\wt \beta^{(i)}$:
\begin{align}
    \beta^{(i)} = \beta_0 + \wt \beta^{(i)}, \text{ for every } i=1,2,\cdots, t, \label{eq_re_mt}
where every entry of $\wt \beta^{(i)}$ is drawn i.i.d. from a Gaussian distribution with zero mean and ${\mu^2}/p$ variance.
Thus, $\mu$ is a parameter for modeling the difference between different tasks.
We study two fundamental questions in this setting.
First, what should the width ($r$) of the feature layer $B$ be?
Second, when does HPS transfer positively to a particular task, depending the sample size $n$ and the (model shift) parameter $\mu$?
Since all tasks are symmetric in this setting, we analyze the average (limiting) risk
\[ g_r(n, \mu) = \frac {1}{t} \sum_{i=1}^t \big(L_i(B^{\star} a_i^{\star}) + \frac{p \sigma^2}{n - p} \cdot \norm{a_i^{\star}}^2\big). \]
We address both questions by characterizing $g_r(n, \mu)$, which accurately approximates $\frac{1}{t} \sum_{i=1}^t L_i(\hat{\beta}_i^{\MTL})$ according to Theorem \ref{thm_many_tasks}.
Recall from Lemma \ref{lem_hat_v} that the excess risk of the OLS estimator satisfies that $L_i(\hat{\beta}_i^{\STL}) = \frac{\sigma^2 p}{n - p} + o_p(1)$, in which $o_p(1)$ vanishes to zero as $p$ approaches infinity.
We compare $g_r(n, \mu)$ with the OLS estimate in the following claim.
\begin{claim}[Choosing $r$ for HPS in the random-effect model]\label{claim_re_multi}
    Suppose the multi-task setting with random-effect linear models under equation \eqref{eq_mtl_data} and \eqref{eq_re_mt} holds.
    Let $\epsilon = o_p(\norm{\beta_0}^2 + \mu^2)$ be a small error term that decreases to zero as $p$ goes to infinity.
    Let $r \in [1, r]$.
    W.h.p. over the randomness of $B^{\star}$, the following holds:
    \begin{enumerate}
        \item[i)] If $\mu^2 \ge \frac{\sigma^2 p^2}{(n - p) \tr[\Sigma]}$, then $g_r(n, \mu) \ge \frac{\sigma^2 p}{n - p} + \epsilon$, for any $1\le r < t$.
    	\item[i)] If $\mu^2 < \frac{\sigma^2 p^2}{(n - p)\tr[\Sigma]} - \frac{t p}{\bigtr{\Sigma}} \epsilon$, then $g_r(n, \mu)$ is minimized when $r=1$.
    	Additionally, $g_1(n, \mu) \le \frac{\sigma^2 p}{n - p} + \epsilon$.

@@

    We first simplify the expression of $g_r(n, \mu)$ using its definition
    \begin{align*}
        g_r(n, \mu) &= \frac 1 t \sum_{i=1}^t \big(L_i(B^{\star} a_i^{\star}) + \frac{p\sigma^2}{n - p} \cdot \norm{a_i^{\star}}^2\big) \\
        &= \Big(\frac 1 t \sum_{i=1}^t \bignorm{\Sigma^{1/2} (B^{\star} a_i^{\star} - \beta^{(i)})}^2 \Big)
        + \Big(\frac 1 t \sum_{i=1}^t \frac{p\sigma^2}{n - p} \cdot \norm{a_i^{\star}}^2 \Big) \\
        &= \frac 1 t \bignormFro{\Sigma^{1/2} (B^{\star} A^{\star} - B^{\star})}^2 + \frac{r}{t} \cdot \frac{p\sigma^2}{n - p}.
    \end{align*}
    In the last step, we use the matrix notation to write the sum for the first part.
    We use the fact that $\sum_{i=1}^t \norm{a_i^{\star}}^2 = r$ because $A^{\star} = [a_1^{\star}, \dots, a_t^{\star}]$ satisfies that ${A^{\star}}^{\top} A^{\star} = \id_{p\times p}$ following its definition in equation \eqref{eq_A_star}.
    Using this condition on $A^{\star}$, we further get that
    \begin{align*}
        \bignormFro{\Sigma^{1/2}(B^{\star} A^{\star} - B^{\star})}^2
        = \bigtr{{B^{\star}}^{\top} \Sigma B^{\star} (\id_{p \times p} - A^{\star}{A^{\star}}^{\top})},
    \end{align*}
    which is precisely equal to the smallest $(t - r)$ eigenvalues of ${B^{\star}}^{\top}\Sigma B^{\star}$.
    Using the concentration of Gaussian random vectors in Lemma \ref{largedeviation}, w.h.p. the $(i, j)$-th entry of ${B^{\star}}^{\top} \Sigma B^{\star}$ is equal to

@@

    	{\beta^{(i)}}^{\top}\Sigma  {\beta^{(j)}} =\beta_0^\top \Sigma \beta_0 + \delta_{i,j} \frac{\mu^2 }{p}\bigtr{\Sigma} + \OO\left(p^{-1/2+c}\|\beta_0\|^2+ p^{-1/2+c} \mu^2\right)

@@

    for any constant $c>0$, where $\delta_{i,j} = 1$ if and only if $i = j$.
    Thus, ${B^{\star}}^{\top} \Sigma B^{\star}$ is equal to a rank-$1$ matrix with (spectral) norm $t\cdot \beta_0^{\top} \Sigma \beta_0$ plus Gaussian perturbation.
    Using the concentration result of Lemma \ref{largedeviation} \todo{check}, we have that
    \begin{align*}
        \lambda_1 &= \Big(1+\OO(p^{-1/2+c})\Big)\cdot\left(t\cdot \beta_0^\top \Sigma \beta_0  + \frac{\mu^2}{ p}\bigtr{\Sigma}\right) , \text{ and } \\
        \lambda_i &= \Big(1+\OO(p^{-1/2+c})\Big)\cdot \frac{\mu^2}{ p}\bigtr{\Sigma}, \text{ for } i=2,\cdots, t.
    \end{align*}
    Thus, w.h.p. the sum of smallest $(t - r)$ eigenvalues of ${B^{\star}}^{\top}\Sigma  B^{\star}$ is equal to
    $(1+\OO(p^{-1/2+c}))\cdot(t-r)\frac{\mu^2 \bigtr{\Sigma}}{p}$.
    Let $\epsilon = o_p(1)$ denote the error term.

@@

    We conclude that w.h.p.,
    \begin{align}
        g_r(n, \mu) - \frac{\sigma^2 p}{n - p} &= \bigbrace{1 - \frac{r}{t}} \frac{\mu^2}{p}\bigtr{\Sigma} + \frac{r}{t} \cdot \frac{p\sigma^2 }{n - p} + \epsilon - \frac{\sigma^2 p}{n - p} \nonumber \\
        &= \Big(1 - \frac r t\Big) \cdot \Big(\frac{\mu^2\bigtr{\Sigma}}{p} - \frac{p\sigma^2}{n - p}\Big) + \epsilon. \label{eq_re_final}
    Now we are ready to finish the proof.
    For claim i), if $\mu^2 \ge \frac{\sigma^2 p^2}{(n - p) \bigtr{\Sigma}}$, the coefficient of $(1 - \frac r t)$ in equation \eqref{eq_re_final} is non-negative, and claim i) follows.
    For claim ii), if $\mu^2 < \frac{\sigma^2 p^2}{(n - p)\bigtr{\Sigma}} - \frac{p t}{\bigtr{\Sigma}}\epsilon$, the coefficient of $(1 - \frac r t)$ in equation \eqref{eq_re_final} is negative.
    Furthermore, equation \eqref{eq_re_final} is minimized when $r = 1$, and claim ii) follows since the RHS of equation \eqref{eq_re_final} is at most $\epsilon$.

@@

\begin{figure}[!t]
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/multitask_width.eps}
		\caption{Varying width}
		\label{fig_sec4_width}
	\end{subfigure}
	\begin{subfigure}[b]{0.5\textwidth}
		\centering
		\includegraphics[width=0.8\textwidth]{figures/multitask_transfer.eps}
		\caption{Varying sample size and model shift parameter}
		\label{fig_sec4_transfer}
	\end{subfigure}
	\caption{We illustrate the performance of HPS in the random-effect model with $p = 100, t = 10, \sigma = 1/2$.
	Figure \ref{fig_sec4_width} shows that from $1$ to $t = 10$, setting width $r = 1$ gives the lowest excess risk for task $t$.
	The result is averaged over three random seeds.
	This simulation uses $\mu = 0.05$.
	Figure \ref{fig_sec4_transfer} fixes $r = 1$ while varying $\mu$ and $n$. We observe that depending on $\mu$ and $n$, HPS may provide a positive transfer or a negative transfer to task $t$ (similar to Figure \ref{fig_sec3_model_shift}). The precise condition for deciding which one is provided in Claim \ref{claim_re_multi}.}
	\label{fig_sec4}
\end{figure}
